### 2011-02-10 paper/introduction.tex
intro edits

\acmFig{robot_close.jpg}{1}{The quadruped robot. The translucent parts were created with 3D-printing technology. Videos of the gaits we evolved for this robot can be viewed online~\cite{quadratotWeb}.}

@@

Once the design goals are decided upon, gaits may be obtained by one of several methods.  Gaits may be designed manually by an expert, as for Asimo and Big Dog \cite{chestnutt2006footstep, raibert2008bigdog}.  Alternately, they may be learned through repeated trial and evaluation.  Learned gaits offer several advantages over manually designed gaits.  Automatically learning gaits can save valuable engineering time and can allow customization of gaits to a particular robot and its unique actuators.  Most importantly, in some cases learned gaits can outperform manually designed gaits~\cite{hornby2005autonomous, valsalam2008modular}.

In this paper we compare the performance of three different methods of designing gaits: manual design, parameterized gaits optimized with seven different learning methods, and gaits generated by evolving neural networks with the HyperNEAT generative encoding~\cite{stanley2009hypercube}. While some of these methods, such as HyperNEAT, have been tested in simulation~\cite{clune2009evolving, clune2011performance}, we investigate how they perform when evolving on a physical robot (\figref{robot_close.jpg}).

Previous work has shown that quadruped gaits perform better when they are \emph{regular} (i.e.\ when the legs are coordinated)~\cite{clune2009evolving, clune2011performance,valsalam2008modular}. For example, HyperNEAT produced fast, natural gaits in part because its bias towards regular gaits created coordinated movements that outperformed gaits evolved by an encoding not biased towards regularity~\cite{clune2009evolving, clune2011performance}. One of the motivations of this paper is to investigate whether any learning method biased towards regularity would perform well at producing quadruped gaits, or whether HyperNEAT's high performance is due to additional factors, such as its abstraction of biological development (described below). We test this hypothesis by comparing HyperNEAT to seven machine learning algorithms biased toward regularity.

A final motivation of this problem is to simply evolve effective gaits for a physical robot. Because HyperNEAT performed well in simulation, it is interesting to test whether it can produce a fast gait for a physical robot. It is additionally interesting to test how more traditional machine learning techniques compete with evolutionary algorithms when evolving in hardware.

>>>>

\acmFig{robot_close.jpg}{1}{The quadruped robot. The translucent parts
  were created with 3D-printing technology. Videos of the gaits
  evolved for this robot can be viewed online~\cite{quadratotWeb}.}

@@

Once the design goals are decided upon, gaits may be obtained by one
of several methods.  Gaits may be designed manually by an expert, as
for Asimo and Big Dog \cite{chestnutt2006footstep, raibert2008bigdog}.
Alternately, they may be learned through repeated trial and
evaluation.  Learned gaits offer several advantages over manually
designed gaits.  Automatically learning gaits can save valuable
engineering time and can allow customization of gaits to a particular
robot and its unique actuators.  Most importantly, in many cases
learned gaits can outperform manually designed
gaits~\cite{hornby2005autonomous, valsalam2008modular}.

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\cite{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\cite{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_close.jpg}).

Previous work has shown that quadruped gaits perform better when they
are \emph{regular} (i.e.\ when the legs are
coordinated)~\cite{clune2009evolving,
  clune2011performance,valsalam2008modular}. For example, HyperNEAT
produced fast, natural gaits in part because its bias towards regular
gaits created coordinated movements that outperformed gaits evolved by
an encoding not biased towards regularity~\cite{clune2009evolving,
  clune2011performance}. One of the motivations of this paper is to
investigate whether any learning method biased towards regularity
would perform well at producing quadruped gaits, or whether
HyperNEAT's high performance is due to additional factors, such as its
abstraction of biological development (described below). We test this
hypothesis by comparing HyperNEAT to six local search algorithms with
a parametrization biased toward regularity.

A final motivation of this problem is to simply evolve effective gaits
for a physical robot. Because HyperNEAT performed well in simulation,
it is interesting to test whether it can produce a fast gait for a
physical robot. It is additionally interesting to test how more
traditional gait optimization techniques compete with evolutionary
algorithms when evolving in hardware.
